---
title: "Bootstrap analyses"
author: "Peter Kamerman and Antonia Wadley"
date: "7 July 2016"
output: 
    github_document:
        toc: true
        toc_depth: 4
---

**Last updated: `r format(Sys.time(), "%B %d, %Y")`**

****

# Session setup
```{r sessionSetup, message = FALSE}
# Load packages
library(svglite)
library(readr)
library(dplyr)
library(tidyr)
library(knitr)

# set seed
set.seed(1234)

# knitr chunk options
opts_chunk$set(echo = TRUE,
               warning = FALSE,
               message = FALSE,
               cache = TRUE,
               cache.extra = rand_seed,
               fig.path = './figures/',
               fig.width = 7,
               fig.height = 7,
               dev = c('png', 'svg'),
               tidy = FALSE)
```

# Load data
```{r import_data}
data <- read_csv('./data/data.csv')
data_activity <- read_csv('./data/activity.data.csv')
```

# Quick look
```{r quick_look}
############################################################
#                                                          #
#                           Data                           #
#                                                          #
############################################################
dim(data)
names(data)
head(data)
tail(data)
glimpse(data)
# summary(data)

############################################################
#                                                          #
#                      Activity data                       #
#                                                          #
############################################################
dim(data_activity)
names(data_activity)
head(data_activity)
tail(data_activity)
glimpse(data_activity)
# summary(data_activity)
```

# Process data
```{r process_data}
############################################################
#                                                          #
#               Merge data and data_activity               #
#                                                          #
############################################################
data <- data %>%
    left_join(data_activity, by = 'id') %>%
    mutate(pain = pain.x) %>%
    select(-pain.x, -pain.y)

############################################################
#                                                          #
#                      Column formats                      #
#                                                          #
############################################################
# Convert chr values to factors
data_1 <- data %>%
    mutate(pain = factor(pain),
           actigraphy = factor(actigraphy),
           eq5d_vas = eq5d_vas/100)

# Convert eq5d mobility to numeric so that 
# Spearman's correlation doesn't throw a fit
data_1$eq5d_mobility[data_1$eq5d_mobility == 'no_problems'] <- 0 
data_1$eq5d_mobility[data_1$eq5d_mobility == 'some_problems'] <- 1 
data_1$eq5d_mobility <- as.numeric(data_1$eq5d_mobility)

############################################################
#                                                          #
#              Generate subsets for use later              #
#                                                          #
############################################################
# Pain only
data_pain <- data_1 %>%
    filter(pain == 'Pain')
# Activity only
data_activity <- data_1 %>%
    filter(actigraphy == 'Yes')
## Activity + pain
data_act_pain <- data_activity %>%
    filter(pain == 'Pain')
## Activity + no pain
data_act_no.pain <- data_activity %>%
    filter(pain == 'No pain')
```

****

# Spearman correlation

## Spearman bootstrapping and permutation function
```{r spearmans}
############################################################
#                                                          #
#  Bootstrap Spearman's correlation with permuted p-value  #
#                                                          #
############################################################

# Create function to generate a bootstrap Spearman Rho
# and to permute a p-value. 

# 10000 iterations for both tests
# Random seed = 1234

# Arguments
## data:    Dataframe containing the data to be analysed.
## var_?:   Name of the two variables to analyse. Use the form: 'name'.

boot_spear <- function(data, var_1, var_2) {
    
    # Set seed
    set.seed(1234)
    
    # Create vector of variable names
    variable_names <- c(var_1, var_2)
    
    # Subset input data according to variable names 
    # and filter out incomplete cases
    
    # Attach required package
    require(dplyr)
    
    data_subset <- data[ , names(data) %in% variable_names] %>%
        filter(complete.cases(.)) %>%
        select_(var_1, var_2) # use standard evaluation
    
    # Check the columns are correct
    if(var_1 %in% names(data_pain) != TRUE |
       var_2 %in% names(data_pain) != TRUE) {
               warning('Column names of subsetted data DO NOT match the variable names provided in \'var_1\' and \'var_2\'')
        } 
    
    ###################
    #  Bootstrapping  #
    ###################
    
    # Attach required packages
    require(boot)
    
    # Boot 'statistic' function
    spear <- function(data_subset, i) {
        d <- data_subset[i, ]
        cor(d[[1]], d[[2]],
            method = 'spearman')
    }
    
    # Generate bootstrap sample
    boot_output <- boot(data_subset, 
                        statistic = spear,
                        R = 10000,
                        sim = 'ordinary',
                        stype = 'i')
    
    # Get 95%CI of bootstrapped sample
    boot_95CI <- boot.ci(boot_output, type = 'basic')
    
    # Generate plot of bootstrapped sample 
    boot_plot <- plot(boot_output)

    #########################
    #  p-value permutation  #
    #########################
    
    # Attach required packages
    require(coin)
    
    # Generate permuted p-value
    perm_output <- spearman_test(data_subset[[1]] ~ data_subset[[2]],
                                 distribution = 'approximate',
                                 B = 10000)
    
    ###########################
    #  Bring it all together  #
    ###########################
    
    # Create list of objects generated so far
    output <- list(paste0('INPUT VARIABLES: ', 
                          variable_names[1], ' and ',
                          variable_names[2]),
                   paste0('Passed internal subsetting check'),
                   paste0('ORIGINAL t0 = ', 
                          round(mean(boot_output$t), 7)),
                   boot_plot,
                   boot_95CI,
                   perm_output)
    
    # Print list
    output
}
```

## Analysis

### WHOLE COHORT

#### EQ5D VAS vs resilience
```{r eq5dvas_res}

boot_spear(data = data_1, 
           var_1 = 'eq5d_vas', 
           var_2 = 'resilience')

```

### PAIN COHORT ONLY
#### Worst pain vs resilience
```{r wpi_res}

boot_spear(data = data_pain, 
           var_1 = 'wpi', 
           var_2 = 'resilience')

```

#### Least pain vs resilience
```{r lpi_res}

boot_spear(data = data_pain, 
           var_1 = 'lpi', 
           var_2 = 'resilience')

```

#### Nadir CD4 vs resilience
```{r cd4_res}

boot_spear(data = data_pain, 
           var_1 = 'nadir_cd4', 
           var_2 = 'resilience')

```

#### BPI inteference vs resilience
```{r bpi_res}

boot_spear(data = data_pain, 
           var_1 = 'bpi_i', 
           var_2 = 'resilience')

```

#### BPI inteference vs worst pain
```{r bpi_wpi}

boot_spear(data = data_pain, 
           var_1 = 'bpi_i', 
           var_2 = 'wpi')

```

#### EQ5D VAS vs worst pain
```{r eq5dv_wpi}

boot_spear(data = data_pain, 
           var_1 = 'eq5d_vas', 
           var_2 = 'wpi')

```

#### EQ5D mobility vs BPI walking
```{r eq5dm_bpiw}

boot_spear(data = data_pain, 
           var_1 = 'eq5d_mobility', 
           var_2 = 'bpi_walking')

```

### ACTIGRAPHY COHORT 
_(pain and no pain)_

#### Median daily activity vs resilience
```{r act_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'median_activity', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'median_activity', 
           var_2 = 'resilience')

```

#### Minutes active per day vs resilience
```{r act2_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'active_min', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'active_min', 
           var_2 = 'resilience')

```

#### Time spent at 0% of maximum activity vs resilience
```{r q0_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'min_q0', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'min_q0', 
           var_2 = 'resilience')
```

#### Time spent in the first quartile of maximum activity vs resilience
```{r q1_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'min_q1', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'min_q1', 
           var_2 = 'resilience')
```

#### Time spent in the second quartile of maximum activity vs resilience
```{r q2_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'min_q2', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'min_q2', 
           var_2 = 'resilience')
```

#### Time spent in the third quartile of maximum activity vs resilience
```{r q3_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'min_q3', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'min_q3', 
           var_2 = 'resilience')
```

#### Time spent in the fourth quartile of maximum activity vs resilience
```{r q4_res}

############################################################
#                                                          #
#                           Pain                           #
#                                                          #
############################################################
boot_spear(data = data_act_pain, 
           var_1 = 'min_q4', 
           var_2 = 'resilience')

############################################################
#                                                          #
#                          No Pain                         #
#                                                          #
############################################################
boot_spear(data = data_act_no.pain, 
           var_1 = 'min_q4', 
           var_2 = 'resilience')
```

### ACTIGRAPHY COHORT
_(pain sub-group only)_

#### BPI walking vs median daily activity
```{r bpi_act}

boot_spear(data = data_act_pain, 
           var_1 = 'bpi_walking', 
           var_2 = 'median_activity')

```

#### BPI walking vs minutes active per day
```{r bpi_act2}

boot_spear(data = data_act_pain, 
           var_1 = 'bpi_walking', 
           var_2 = 'active_min')

```

#### EQ5D mobility vs median daily activity
```{r eq5dm_act}

boot_spear(data = data_act_pain, 
           var_1 = 'eq5d_mobility', 
           var_2 = 'median_activity')

```

#### EQ5D mobility vs minutes active per day
```{r eq5dm_act2}

boot_spear(data = data_act_pain, 
           var_1 = 'eq5d_mobility', 
           var_2 = 'active_min')

```

#### Worst pain vs median daily activity
```{r wpi_act}

boot_spear(data = data_act_pain, 
           var_1 = 'wpi', 
           var_2 = 'median_activity')

```

#### Least pain vs median daily activity
```{r lpi_act}

boot_spear(data = data_act_pain, 
           var_1 = 'lpi', 
           var_2 = 'median_activity')

```

****

# Wilcoxon rank sum test

## Wilcoxon bootstrapping and permutation function
```{r wilcoxon}
############################################################
#                                                          #
#      Bootstrap Wilcoxon test with permuted p-value       #
#                                                          #
############################################################

# Create function to generate a bootstrap Wilcoxon Zeta
# and to permute a p-value. 

# 10000 iterations for both tests
# Random seed = 1234

# Arguments
## data:    Dataframe containing the data to be analysed.
## var_?:   Name of the two variables to analyse. Use the form: 'name'.

boot_wilcox <- function(data, var_1, var_2) {
    
    # Set seed
    set.seed(1234)
    
    # Create vector of variable names
    variable_names <- c(var_1, var_2)
    
    # Subset input data according to variable names,  
    # filter out incomplete cases, and order the columns
    
    # Attach required package
    require(dplyr)
    
    data_subset <- data[ , names(data) %in% variable_names] %>%
        filter(complete.cases(.)) %>%
        select_(var_1, var_2) # use standard evaluation
    
    # Check the columns are correct
    if(var_1 %in% names(data_pain) != TRUE |
       var_2 %in% names(data_pain) != TRUE) {
               warning('Column names of subsetted data DO NOT match the variable names provided in \'var_1\' and \'var_2\'')
        } 
    
    ###################
    #  Bootstrapping  #
    ###################
    
    # Attach required packages
    require(boot)
    
    # Boot 'statistic' function
    wilcox <- function(data_subset, i) {
        d <- data_subset[i, ]
        foo <- wilcox_test(d[[1]] ~ d[[2]])
        foo@statistic@teststatistic
    }
    
    # Generate bootstrap sample
    boot_output <- boot(data_subset, 
                        statistic = wilcox,
                        R = 10000,
                        sim = 'ordinary',
                        stype = 'i',
                        parallel = 'multicore')
    
    # Get 95%CI of bootstrapped sample
    boot_95CI <- boot.ci(boot_output, type = 'basic')
    
    # Generate plot of bootstrapped sample 
    boot_plot <- plot(boot_output)

    #########################
    #  p-value permutation  #
    #########################
    
    # Attach required packages
    require(coin)
    
    # Generate permuted p-value
    perm_output <- wilcox_test(data_subset[[1]] ~
                                   data_subset[[2]],
                               distribution = 'approximate',
                               B = 10000)
    
    ###########################
    #  Bring it all together  #
    ###########################
    
    # Create list of objects generated so far
    output <- list(paste0('INPUT VARIABLES: ', 
                          variable_names[1], ' and ',
                          variable_names[2]),
                   paste0('Passed internal subsetting check'),
                   paste0('ORIGINAL t0 = ', 
                          round(mean(boot_output$t), 7)),
                   boot_plot,
                   boot_95CI,
                   perm_output)
    
    # Print list
    output
}
```

## Analysis

### ACTIGRAPHY
_(whole cohort)_

#### Minutes active per day vs pain/no pain
```{r act2_pain}

boot_wilcox(data = data_activity,
            var_1 = 'active_min',
            var_2 = 'pain')

```

#### Median daily active vs pain/no pain
```{r act_pain}

boot_wilcox(data = data_activity,
            var_1 = 'median_activity',
            var_2 = 'pain')

```

#### Time spent at 0% maximum activity vs pain/no pain
```{r q0_pain}

boot_wilcox(data = data_activity,
            var_1 = 'min_q0',
            var_2 = 'pain')

```

#### Time spent in first quartile of maximum activity vs pain/no pain
```{r q1_pain}

boot_wilcox(data = data_activity,
            var_1 = 'min_q1',
            var_2 = 'pain')

```

#### Time spent in second quartile of maximum activity vs pain/no pain
```{r q2_pain}

boot_wilcox(data = data_activity,
            var_1 = 'min_q2',
            var_2 = 'pain')

```

#### Time spent in third quartile of maximum activity vs pain/no pain
```{r q3_pain}

boot_wilcox(data = data_activity,
            var_1 = 'min_q3',
            var_2 = 'pain')

```

#### Time spent in fourth quartile of maximum activity vs pain/no pain
```{r q4_pain}

boot_wilcox(data = data_activity,
            var_1 = 'min_q4',
            var_2 = 'pain')

```

****

# Session information
```{r session_information}

sessionInfo()

```