---
title: 'Supplementary plots'
author: 'Peter Kamerman'
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
    md_document:
        variant: markdown_github
---
## Load required packages and set chunk options
```{r LoadPackages}
# Load packages
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(grid)
library(cowplot)

# Load palette
cb8.categorical <- c("#0072B2", "#D55E00", "#009E73", "#F0E442", "#56B4E9", "#E69F00", "#CC79A7", "#999999")

# knitr chunk options
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               fig.path = './figures/',
               fig.width = 11.7,
               fig.height = 8.2,
               dev = c('png', 'pdf'),
               tidy = TRUE, 
               tidy.opts = list(width.cutoff = 65))
```


## Resilience vs Activity dot-plots 
_**68 participants, 37 with pain and 31 without pain**_
### Import data
```{r DataImport}
data.res <- read_csv('./data/resilience.V.endurance.csv', col_names = TRUE)
```

### Quick look
```{r DataProcessing}
# Top-and-tail
head(data.res)
tail(data.res)
glimpse(data.res)

# Process data
data.res.1 <- data.res %>%
    mutate(pain = factor(Pain, 
                         levels = c('no.pain', 'pain'),
                         labels = c('No pain', 'Pain'))) %>%
    select(-ID, -Pain) %>%
    rename(activity = Median_activity,
           rs.score = RS_corrected, 
           zero = Minutes_at_0_activity,
           q1 = Minutes_at_q1,
           q2 = Minutes_at_q2,
           q3 = Minutes_at_q3,
           q4 = Minutes_at_q4) 
glimpse(data.res.1)
```

## Create resilience plots
```{r Plot}
# Plot_1: resilience score vs median activity
Resilience.activity <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = activity,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 5, shape = 21, alpha = 0.7) +
    labs(x = '\nResilience Scale score',
         y = 'Median activity (counts)\n',
         title = 'Resilience Scale score vs median activity\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
Resilience.activity

# Plot_2: resilience score vs time spent at 0% of max activity
RS.EnduranceZero <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = zero,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    labs(title = 'Time spent at 0% of maximum activity\n',
         x = '\nResilience Scale score',
         y = 'Minutes per day\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
RS.EnduranceZero

# Plot_3: resilience score vs time spent at 1-24% of max activity
RS.EnduranceQ1 <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = q1,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    labs(title = 'Time spent at 1-24% of maximum activity\n',
         x = '\nResilience Scale score',
         y = 'Minutes per day\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
RS.EnduranceQ1

# Plot_4: resilience score vs time spent at 25-49% of max activity
RS.EnduranceQ2 <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = q2,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    labs(title = 'Time spent at 25-49% of maximum activity\n',
         x = '\nResilience Scale score',
         y = 'Minutes per day\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
   theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
RS.EnduranceQ2

# Plot_5: resilience score vs time spent at 50-74% of max activity
RS.EnduranceQ3 <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = q3,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    labs(title = 'Time spent at 50-74% of maximum activity\n',
         x = '\nResilience Scale score',
         y = 'Minutes per day\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
RS.EnduranceQ3

# Plot 6: resilience score vs time spent at >75% of max activity
RS.EnduranceQ4 <- ggplot(data.res.1,
                 aes(x = rs.score,
                     y = q4,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'lm') +
    labs(title = 'Time spent at >75% of maximum activity\n',
         x = '\nResilience Scale score',
         y = 'Minutes per day\n') +
    facet_grid(.~pain) +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none',
          panel.margin.x = unit(1, 'lines'))
RS.EnduranceQ4
```

## Brief Pain Inventory vs activity dot-plots
_**37 participants with pain, only**_
### Import data
```{r DataImport.bpi}
data.bpi <- read_csv('./data/bpi.activity.csv', col_names = TRUE)
```

### Quick look
```{r DataProcessing.bpi}
# Top-and-tail
head(data.bpi)
tail(data.bpi)
glimpse(data.bpi)

# Process data
data.bpi.1 <- data.bpi %>%
    mutate(pain = as.factor(Pain)) %>%
    select(-ID, -Pain) %>%
    rename(activity = Median_activity,
           bpi_interference = BPI_interference, 
           worst_pain = Worst_pain,
           least_pain = Least_pain) 
glimpse(data.bpi.1)
```

## Create bpi plots
```{r Plot.bpi}
# Plot_7: BPI interference vs median activity
Interference.activity <- ggplot(data.bpi.1,
                 aes(x = bpi_interference,
                     y = activity,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    labs(x = '\nPain inteference (11-point NRS)',
         y = 'Median activity (counts)\n',
         title = 'Pain interference in the last week vs activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none')
Interference.activity

# Plot_8: Worst pain vs median activity
WorstPain.activity <- ggplot(data.bpi.1,
                 aes(x = worst_pain,
                     y = activity,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    labs(x = '\nPain intensity (11-point NRS)',
         y = 'Median activity (counts)\n',
         title = 'Worst pain in the last week vs activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none')
WorstPain.activity

# Plot_9: Least pain vs median activity
LeastPain.activity <- ggplot(data.bpi.1,
                 aes(x = least_pain,
                     y = activity,
                     fill = pain,
                     colour = pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    labs(x = '\nPain intensity (11-point NRS)',
         y = 'Median activity (counts)\n',
         title = 'Least pain in the last week vs activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none')
LeastPain.activity
```

## Activity vs predictor scatterplots
_**68 participants with (37) and without (31) pain**_
### Import data
```{r DataImport.act}
data.act <- read_csv('./data/activity.V.predictors.csv', col_names = TRUE)
```

### Quick look
```{r DataProcessing.act}
# Top-and-tail
head(data.act)
tail(data.act)
glimpse(data.act)

# Process data
data.act.1 <- data.act %>%
    mutate(ID = factor(ID),
           pain = factor(pain),
           sex = factor(sex),
           education = factor(education, 
                              levels = c('No education', 
                                         'Completed grade 7', 
                                         'Completed grade 12', 
                                         'Tertiary education'), 
                                         ordered = TRUE),
           employed = factor(employed)) %>%
    group_by(ID) %>%
    gather(key = worry_about, value = rating, money:health) %>%
    mutate(rating = factor(rating, 
                           levels = c('Not at all',
                                         'Rarely',
                                         'Sometimes', 
                                         'Often', 
                                         'Nearly all the time'), 
                                         ordered = TRUE)) %>%
    arrange(ID)

# Quick look
str(data.act.1)
```

## Create bpi plots
```{r Plot.act}
# Plot_10: median activity vs. presence of pain
Pain.activity <- ggplot(data.act.1,
                 aes(x = pain,
                     y = median_activity,
                     fill = pain,
                     colour = pain)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    labs(y = 'Median activity (counts)\n',
         title = 'Median weekly activity by the presence of pain\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.title.x = element_blank(),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none')
Pain.activity

# Plot_11: median activity vs age
Age.activity <- ggplot(data.act.1,
                 aes(x = age,
                     y = median_activity)) +
    geom_jitter(aes(fill = pain, colour = pain),
                size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'loess') +
    labs(x = '\nAge (years)',
         y = 'Median activity (counts)\n',
         title = 'Median weekly activity vs. Age\n(loess smoother)') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'top',
          legend.text = element_text(size = 16),
          legend.title = element_blank())
Age.activity

# Plot_12: median activity vs. sex
Sex.activity <- ggplot(data.act.1,
                 aes(x = sex,
                     y = median_activity,
                     fill = sex,
                     colour = sex)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    labs(y = 'Median activity (counts)\n',
         title = 'Median weekly activity by sex\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.title.x = element_blank(),
          panel.grid.major = element_line(colour = 'gray90'),
          panel.background = element_blank(),
          legend.position = 'none')
Sex.activity

# Plot_13: median activity vs. education
## Need to summarise data and use stat = 'identity'
edu <- data.act.1 %>%
    filter(!is.na(education)) %>%
    group_by(education) %>%
    summarise(Median = median(median_activity), 
              Min = min(median_activity),
              q25 = quantile(median_activity, 0.25),
              q75 = quantile(median_activity, 0.75),
              Max = max(median_activity))

Edu.activity <- ggplot(edu,
                 aes(x = education,
                     fill = education,
                     colour = education)) +
    geom_errorbar(aes(ymin = Min, ymax = Max), 
                  size = 1,
                  width = 0) +
    geom_crossbar(aes(y = Median, ymin = q25, ymax = q75), 
             fill = 'white', 
             colour = 'white') +
    geom_crossbar(aes(y = Median, ymin = q25, ymax = q75),
             size = 1, 
             alpha = 0.5) +
    labs(y = 'Median activity (counts)\n',
         title = 'Median weekly activity by education level\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.text.x = element_text(angle = 30, hjust = 1),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          axis.title.x = element_blank(),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'none')
Edu.activity

# Plot_14: Median weekly activity vs. BMI
BMI.activity <- ggplot(data.act.1,
                 aes(x = bmi,
                     y = median_activity)) +
    geom_jitter(aes(fill = pain, colour = pain),
                size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'loess') +
    labs(x = '\nBMI (m/kg^2)',
         y = 'Median activity (counts)\n',
         title = 'Median weekly activity vs. BMI\n(loess smoother)') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'top',
          legend.text = element_text(size = 16),
          legend.title = element_blank())
BMI.activity

# Plot_15: median activity vs. presence of pain
Employ.activity <- ggplot(data.act.1,
                 aes(x = employed,
                     y = median_activity,
                     fill = employed,
                     colour = employed)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    labs(y = 'Median activity (counts)\n',
         title = 'Median weekly activity by employment status\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          axis.title.x = element_blank(),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'none')
Employ.activity

# Plot_16: Median weekly activity vs. RS score
RS.activity <- ggplot(data.act.1,
                 aes(x = rs_score,
                     y = median_activity)) +
    geom_jitter(aes(fill = pain, colour = pain),
                size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'loess') +
    labs(x = '\nResilience Scale score',
         y = 'Median activity (counts)\n',
         title = 'Median weekly activity vs. RS score\n(loess smoother)') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'top',
          legend.text = element_text(size = 16),
          legend.title = element_blank())
RS.activity

# Plot_17: Median weekly activity vs. EQ5D VAS
EQ5D.activity <- ggplot(data.act.1,
                 aes(x = eq5d_vas,
                     y = median_activity)) +
    geom_jitter(aes(fill = pain, colour = pain),
                size = 4, shape = 21, alpha = 0.5) +
    geom_smooth(method = 'loess') +
    labs(x = '\nEQ5D VAS score',
         y = 'Median activity (counts)\n',
         title = 'Median weekly activity vs. EQ5D VAS score\n(loess smoother)') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'top',
          legend.text = element_text(size = 16),
          legend.title = element_blank())
EQ5D.activity
```
```{r Plot.act.b, fig.width = 8.2, fig.height = 11.7}
# Plot_18: median activity vs. Worries (money & family)
## Subset data
worries.a <- data.act.1 %>%
    filter(worry_about != 'food' | worry_about != 'health') %>%
    mutate(worry_about = factor(worry_about)) %>%
    rename(worries_a = worry_about)
worries.a

Worr.activity <- ggplot(data = worries.a,
                 aes(x = rating,
                     y = median_activity,
                     fill = rating,
                     colour = rating)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    facet_grid(worries_a~.) +
    labs(y = 'Median activity (counts)\n',
         title = 'Median weekly activity by level\nof monetary and family worries\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour='black'),
          axis.title.x = element_blank(),
          panel.background=element_blank(),
          panel.grid.major = element_line(colour = 'grey90'),
          legend.position = 'none',
          panel.margin.y = unit(1, 'lines'))
Worr.activity
```

## Session information
```{r SessionInfo}
sessionInfo()
```
