---
output: github_document
---

# Worries data
## Data summaries
_(raw data can be found in [worries.data.csv](./data/worries.data.csv), and the code sheet is provided [below](#code-sheet))_

```{r sessionSetup, message = FALSE, warning = FALSE, include = FALSE}
############################################################
#                                                          #
#                      Session setup                       #
#                                                          #
############################################################

# Load packages
library(svglite)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(grid)

# knitr chunk options
knitr::opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               fig.path = './figures/',
               fig.height = 8,
               fig.width = 8,
               dev = c('svg', 'png'), 
               tidy = FALSE)
```

## Import data
```{r import_data, echo = TRUE}
data <- read_csv('./data/worries.data.csv')
```

```{r quick_look, eval = FALSE, include = FALSE}
############################################################
#                                                          #
#                        Quick look                        #
#                                                          #
############################################################

dim(data)
names(data)
head(data)
tail(data)
glimpse(data)
# summary(data)

```

## Worries about pain
```{r pain}

data_pain <- data %>%
    select(c(2, 4, 5))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_pain, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_pain

df_pain <- as.data.frame(data_pain) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_pain, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about pain',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))

```

## Worries about fatigue
```{r fatigue}

data_fatigue <- data %>%
    select(c(2, 4, 6))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_fatigue, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_fatigue

df_fatigue <- as.data.frame(data_fatigue) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_fatigue, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about fatigue',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))

```

## Worries about health
```{r health}

data_health <- data %>%
    select(c(2, 4, 7))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_health, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_health

df_health <- as.data.frame(data_health) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_health, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about health',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))
```

## Worries about sleep
```{r sleep}

data_sleep <- data %>%
    select(c(2, 4, 8))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_sleep, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_sleep

df_sleep <- as.data.frame(data_sleep) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_sleep, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about sleep',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))
```

## Worries about money
```{r money}

data_money <- data %>%
    select(c(2, 4, 9))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_money, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_money

df_money <- as.data.frame(data_money) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_money, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about money',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))
```

## Worries about family
```{r family}

data_family <- data %>%
    select(c(2, 4, 10))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_family, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_family

df_family <- as.data.frame(data_family) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_family, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about family',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))

```

## Worries about food
```{r food}

data_food <- data %>%
    select(c(2, 4, 11))  %>%
    mutate(`EMPLOYMENT STATUS` = # recode employment status
               ifelse(employment == 'Full time' | 
                          employment == 'Part time/piece work',
                      'employed', 'unemployed')) %>%
    mutate(RATING = factor(worries_about_food, 
                           levels = c('Not at all', 'Rarely',
                                      'Sometimes', 'Often', 
                                      'Nearly all the time'), 
                           ordered = TRUE)) %>%
    select(1, 3, 4, 5) %>%
    rename(`CHRONIC PAIN:` = chronic_pain) %>%
    group_by(`EMPLOYMENT STATUS`, `CHRONIC PAIN:`, RATING) %>%
    summarise(COUNT = n()) %>%
    xtabs(COUNT ~ RATING + `EMPLOYMENT STATUS` + `CHRONIC PAIN:`,
          data = .) %>%
    ftable(.)

data_food

df_food <- as.data.frame(data_food) %>%
    rename(Rating = RATING,
           Employment_status = EMPLOYMENT.STATUS,
           Pain_status = CHRONIC.PAIN.,
           Count = Freq) %>%
    mutate(Percentage = round((Count / sum(.$Count)) * 100)) %>%
    mutate(Pain_status = ifelse(Pain_status == 'Yes', 
                                'Chronic pain', 'No chronic pain')) %>%
    mutate(Employment_status = ifelse(Employment_status == 'employed', 
                                'Employed', 'Unemployed'))
    
ggplot(df_food, aes(x = Rating, y = Percentage)) +
    geom_bar(stat = 'identity') +
    facet_grid(Employment_status ~ Pain_status) +
    labs(title = 'Worries about food',
         x = '\nRating',
         y = 'Percentage of total cohort (%)\n') +
    theme(plot.title = element_text(size = 18),
          axis.title = element_text(size = 16),
          axis.text = element_text(size = 14),
          axis.text.x = element_text(angle = 30, hjust = 1),
          strip.text = element_text(size = 12))
```

****

## Code sheet

**Note:** _'worries\_about\_pain'_ and _'worries\_about\_sleep'_ were deemed to have been assessed twice (firstly in the EQ5D subscales) and they were subcategories of _'worries\_about\_health'_, and therefore only the global concept _('worries\_about\_health')_ was assessed in the manuscript.

- **ID:** Participant id code

- **chronic_pain:** Had chronic pain (pain on most days for $geq$ 3 months) (Yes / No)	

- **actigraphy:** Took part in the actigraphy component of the study (Yes / No)

- **worries\_about\_pain:** Do you worry about your pain? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)

- **worries\_about\_fatigue:** Do you worry about being fatigued? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)

- **worries\_about\_health:** Do you worry about your health? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)	

- **worries\_about\_sleep:** Do you worry about your sleep? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)

- **worries\_about\_money:** Do you worry about money? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)

- **worries\_about\_family:** Do you worry about your pain? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)

- **worries\_about\_food:** Do you worry about your pain? (rating scale: Not at all, Rarely, Sometimes, Often, Nearly all the time)
