---
title: "Actical data analysis"
author: Peter Kamerman 
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  md_document:
    variant: markdown_github
---
```{r Load_required_packages}
# Load packages
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(scales) # required for ggplot
library(exactRankTests)
library(knitr)

# knitr chunk options
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               fig.path = './figures/',
               dev = c('png', 'pdf'),
               tidy = TRUE, 
               tidy.opts = list(width.cutoff = 65))
```

## Import data
```{r Import_data}
# Import
data <- read.csv("./data/activity.data.csv", header = T)
```

## Convert "data" to dplyr object and get an overview of the data before cleaning
```{r convert_to_dplyr}
# Convert data
data.1 <- tbl_df(data)
# Look at data
str(data.1)
```

## Basic data cleaning
### Set column names
_**(Units and definitions remain as described in the README.md file)**_
```{r rename_columns}
# Set names: id  =  Pt ID; pain  =  Pain; employed  =  Employed; time  =  Time; day  =  Day; count  =  Count; pmax.count  =  % count; pptmax.count  =  pcount.ptmax 
# Pass result to "data.1"
names(data.1) <- c("id", "pain", "employed", "time", "day", "count", "pmax.count", "pptmax.count")
```

### Change $id and $pain to factors
```{r format_to_factors}
data.1$id <- factor(data.1$id)
data.1$pain <- factor(data.1$pain, levels = c(0,1), labels = c("no.pain", "pain"))
# Pass "data.1" to "data.2"
data.2 <-data.1
```

### Add a new column ($date-time)
```{r add_date_time_column}
# Add column and pass to "data.3"
data.3 <- unite(data.2, "date", time, day, sep = " ")
data.3$date <- dmy_hms(data.3$date)
```

### Check the dataframe for duplicates
```{r check_for_duplicates}
dim(data.3) 
# There should be (1440m/day * 7 days * 67 participants) + (1440 * 5 days * 1 participant)
(1440*7*67)+(1440*5*1)
# There is a duplicate, so remove it
data.3 <- distinct(data.3)
```

### Create repeating data/time factor to assist data sorting
```{r create_sorting_factor}
# Create sequence from 1 to 136512 (the number of data rows)
# Pass it to "marker"
marker <- tbl_df(data.frame(factor(rep(seq(from = 1, to = 136512, 1), each = 5))))
names(marker) <- "marker"
str(marker)
# Bind "marker" to "data.3", and convert to dplyr object
# Pass to "data.5"
data.5 <- bind_cols(marker, data.3) 
data.5 <- tbl_df(data.5)
```

## Strip out inactive periods
### Find the average pmax.count for each 5-minute period
```{r max_count}
# Find mean pmax.count for each five-minute period, and pass to "marker.2"
marker.2 <- tapply(data.5$pmax.count, data.5$marker, mean)
# Repeat each marker 5 times (so marker and data.5 have the same number of rows)
# Pass to "marker.3" and then convert to a dplyr object
marker.3 <- rep(marker.2, each = 5)
marker.3 <- tbl_df(data.frame(marker.3))
# Column bind "marker.3" and "data.3", convert to a dplyr object, and pass to "data.6"
data.6 <- tbl_df(bind_cols(marker.3, data.5))
```

### Identify and sort active and inactive periods 
_**(mean pmax.count in a 5-minute period > 2.5%  = > active)**_
```{r active_vs_inactive}
# Create new column $marker.4 containing logical of pmax.count > 2.5%
# Pass to "data.7"
data.7 <- mutate(data.6, marker.4 = data.6$marker.3>2.5)
# label days for the 67 with 7 days of data each by creating vector "days"
days <- c("d1", "d2", "d3", "d4", "d5", "d6", "d7") 
# Repeat "days" 1440 times (minutes per 24 hours)
# Pass to "days.2"
days.2 <- rep(days, each = 1440)
# Repeat "days.2" 67 times for the 67 participants with 7 days data
# Pass to "days.3", and name column
days.3 <- data.frame(rep(days.2, times = 67))
names(days.3) <- "days"
# Repeat process for the one paricuipant with 5 days of data
# Pass to "days.4"
days <- c("d1", "d2", "d3", "d4", "d5")
days.4 <- data.frame(rep(days, each = 1440))
names(days.4) <- "days"
# Row bind "days.4" and "days.3"
# Pass to days.5
days.5 <- tbl_df(rbind(days.3, days.4))
# Column bind "days.5" to "data.7"
data.8 <- tbl_df(bind_cols(days.5, data.7))
# Have a quick look at the new dplyr object
str(data.8)
# Choose the columns 
# Pass to "data.9"
data.9 <- select(data.8, id, date, days, pain, employed, count, pmax.count, pptmax.count, marker.4)
# Rename $marker.4 as $active
data.9 <- rename(data.9, active = marker.4)
# Change "?" for patient 90 to "U"
data.9$employed <- as.character(data.9$employed)
data.9$employed[data.9$employed == "?"] <- "U"
data.9$employed <- factor(data.9$employed)
# Have a quick look at "data.9"
str(data.9)
# Group data for sorting
# Pass to "data.10"
data.10 <- group_by(data.9, id, days, date, pain, employed)
# divide into active and inactive periods according to whether the five-minute epochs had a mean pmax.count > 1%
active <- filter(data.10, active == "TRUE")
inactive <- filter(data.10, active == "FALSE")
```

## Summary of pmax.count data
### 24-hour data (active and inactive periods)
#### Median and mean activity per day
```{r pmax.summary.24h.day}
# Extract median and mean data per participant, per day
med.1 <- data.3 %>%
    separate(date, c("day", "time"), sep = " ") %>%
    select(id, pain, employed, day, pmax.count) %>%
    group_by(id, day, pain, employed) %>%
    summarise(Median = median(pmax.count), Mean = round(mean(pmax.count), 1))
med.1
```

#### Median and mean activity per week
```{r pmax.summary.24h.week}
# Extract median and mean data per participant for the week (or 5 days)
med.2 <- data.3 %>%
    separate(date, c("day", "time"), sep = " ") %>%
    select(id, pain, employed, day, pmax.count) %>%
    group_by(id, pain, employed) %>%
    summarise(Median = median(pmax.count), Mean = round(mean(pmax.count), 1))
```

### Active periods only
#### Median and mean activity per day
```{r pmax.summary.active.day}
# Extract median and mean data per participant, per day
med.3 <- active %>%
    ungroup() %>%
    select(id, pain, employed, days, pmax.count) %>%
    group_by(id, days, pain, employed) %>%
    summarise(Median = round(median(as.numeric(pmax.count)),1), Mean = round(mean(as.numeric(pmax.count)),1))
```

#### Median and mean activity per week
```{r pmax.summary.active.week}
# Extract median and mean data per participant for the week (or 5 days)
med.4 <- active %>%
    ungroup() %>%
    select(id, pain, employed, pmax.count) %>%
    group_by(id, pain, employed) %>%
    summarise(Median = round(median(as.numeric(pmax.count)),1), Mean = round(mean(as.numeric(pmax.count)),1))
```

## Univariate analysis of the active period
### Mean time (in minutes) active per day over the week
#### Process the data
```{r time_active}
# Group and pass to "active.1"
active.1 <- group_by(active, id, days, pain, employed)
# Summarise: Number of 5-minute time periods each person spent active per day
# Pass to "time.active"
time.active <- summarise(active.1, times = n())
# Group and pass to "time.1"
time.1 <- ungroup(time.active)
time.1 <- group_by(time.1, id, pain, employed)
# Summarise the mean time spent active for each person across the 7 (or 5) days
time.2 <- summarise(time.1, data = mean(times))
```

#### Pain vs no-pain
```{r time_active_pain}
# Plot
qplot(x = pain, y = data, data = time.2, geom = "boxplot")
# Wilcoxon rank sum test
with(time.2, wilcox.exact(data~pain, paired = F, exact = T))
# N, Median, IQR, range
foo <- time.2 %>%
    group_by(pain) %>%
    select(pain,data) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Employed vs unemployed
```{r time_active_employment}
# Plot
qplot(x = employed, y = data, data = time.2, geom = "boxplot")
# Wilcoxon rank sum test
with(time.2, wilcox.exact(data~employed, paired = F, exact = T))
# N, Median, IQR, range
foo <- time.2 %>%
    group_by(employed) %>%
    select(employed,data) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Conditional plot of employment and pain status
```{r time_active_pain.employment}
# Plot
qplot(x = pain, y = data, data = time.2, facets = .~employed, color = pain, geom = "boxplot")
# N, Median, IQR, range
foo <- time.2 %>%
    group_by(pain, employed) %>%
    select(pain, employed,data) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

### Percentage of actical maximum (pmax.count, %) active per day over the week
#### Process the data
```{r pmax}
pmax.count <- summarise(active.1, p.max = max(pmax.count))
pmax.count.1 <- ungroup(pmax.count)
pmax.count.1 <- group_by(pmax.count.1, id, pain, employed)
pmax.count.2 <- ungroup(pmax.count.1)
pmax.count.2 <- summarise(pmax.count.1, data = max(p.max))
pmax.count.2 <- ungroup(pmax.count.2)
```

#### Pain vs no-pain
```{r pmax_pain}
# Plot
qplot(x = pain, y = data, data = pmax.count.2, geom = "boxplot")
# Wilcoxon rank sum test
with(pmax.count.2, wilcox.exact(data~pain, paired = F, exact = T))
# N, Median, IQR, range
foo <- pmax.count.2 %>%
    group_by(pain) %>%
    select(pain, data) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Employed vs unemployed
```{r pmax_employment}
# Plot
qplot(x = employed, y = data, data = pmax.count.2, geom = "boxplot")
# Wilcoxon rank sum test
with(pmax.count.2, wilcox.exact(data~employed, paired = F, exact = T))
# N, Median, IQR, range
foo <- pmax.count.2 %>%
    group_by(employed) %>%
    select(employed,data) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Conditional plot of employment and pain status
```{r pmax_pain.employment}
# Plot
qplot(x = pain, y = data, data = pmax.count.2, facets = .~employed, color = pain, geom = "boxplot")
# N, Median, IQR, range
foo <- pmax.count.2 %>%
    group_by(pain, employed) %>%
    select(pain, employed,data) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

### Mean sum of daily activity over the week
#### Process the data
```{r total_activity}
activity.1 <- group_by(active.1, id, days, employed, pain)
activity.1 <- summarise(activity.1, data = sum(pmax.count))
activity.2 <- ungroup(activity.1)
activity.2 <- group_by(activity.2, id, employed, pain)
activity.2 <- summarise(activity.2, data = round(sum(data)/length(days),0))
activity.2 <- ungroup(activity.2)
```

#### Pain vs no pain
```{r total_activity_pain}
# Plot
qplot(x = pain, y = data, data = activity.2, geom = "boxplot")
# Wicoxon rank sum test
with(activity.2, wilcox.exact(data~pain, paired = F, exact = T))
# N, Median, IQR, range
foo <- activity.2 %>%
    group_by(pain) %>%
    select(pain,data) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Employed vs unemployed
```{r total_activity_employment}
# Plot
qplot(x = employed, y = data, data = activity.2, geom = "boxplot")
# Wicoxon rank sum test
with(activity.2, wilcox.exact(data~employed, paired = F, exact = T))
# N, Median, IQR, range
foo <- activity.2 %>%
    group_by(employed) %>%
    select(employed,data) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Conditional plot of employment and pain status
```{r total_activity_pain.employment}
# Plot
qplot(x = pain, y = data, data = activity.2, facets = .~employed, colour = pain, geom = "boxplot")
# N, Median, IQR, range
foo <- activity.2 %>%
    group_by(pain, employed) %>%
    select(pain, employed,data) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(data), 
              IQR = IQR(data), 
              Min = min(data), 
              Max = max(data))
kable(foo, format = "markdown", align = "l", digits = 0)
```

## Univariate analysis of complete data 
_**(active and inactive periods, data.10)**_
### Time spent (minutes) in quartiles ($\geq$ 75%, 50-74%, 25-49%, and 0-24%) of percent patient maximum activity count per day over the week
#### Fourth quartile ($\geq$ 75%)
```{r forth.q}
fourth.q <- group_by(data.10, id, date, pain, employed) #> 75%
fourth.q.1 <- filter(fourth.q, pptmax.count>75)
fourth.q.2 <- ungroup(fourth.q.1)
fourth.q.2 <- group_by(fourth.q.2, id, pain, days, employed)
fourth.q.2 <- summarise(fourth.q.2, time = n())
fourth.q.2 <- group_by(fourth.q.2, id, pain, employed)
day <- data.frame(day = c(rep(7, 67), 5))
fourth.q.3 <- summarise(fourth.q.2, time = sum(time))
fourth.q.3 <- bind_cols(fourth.q.3, day)
fourth.q.3 <- mutate(fourth.q.3, time.2 = round(time/day, 2)) # number of minutes per day spent at > 75% patient max
# Save to csv
fq <- fourth.q.3 %>%
    rename("minutes per week" = time, "days of recording" = day, "minutes per day" = time.2)
write.csv(fq, "../activity.plot/data/time.active.4q.csv", row.names = F)
# Ungroup
fourth.q.3 <- ungroup(fourth.q.3)

# Plot
# Univariate
qplot(x = pain, y = time.2, data = fourth.q.3, geom = "boxplot")
# Conditional (with employment status)
qplot(x = pain, y = time.2, data = fourth.q.3, facets = .~employed, colour = pain, geom = "boxplot")
# Wilcoxon rank sum test
with(fourth.q.3, wilcox.exact(time.2~pain, exact = T, paired = F))
# N, Median, IQR, range
foo <- fourth.q.3 %>%
    group_by(pain) %>%
    select(pain,time.2) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- fourth.q.3 %>%
    group_by(employed) %>%
    select(employed,time.2) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- fourth.q.3 %>%
    group_by(pain, employed) %>%
    select(pain, employed, time.2) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Third quartile (50-74%)
```{r third.q}
third.q <- group_by(data.10, id, date, pain, employed) # 50-74%
third.q.1 <- filter(third.q, pptmax.count<= 75 & pptmax.count>50)
third.q.2 <- ungroup(third.q.1)
third.q.2 <- group_by(third.q.2, id, pain, days, employed)
third.q.2 <- summarise(third.q.2, time = n())
third.q.2 <- group_by(third.q.2, id, pain, employed)
third.q.3 <- summarise(third.q.2, time = sum(time))
day <- tbl_df(data.frame(id.1 = as.numeric(as.character(fourth.q.3$id)), days = c(rep(7, 67), 5)))
id <- tbl_df(data.frame(id.1 = as.numeric(as.character(third.q.3$id))))
day.2 <- semi_join(day,id, by = "id.1")
third.q.3 <- bind_cols(third.q.3, day.2)
third.q.3 <- select(third.q.3, id, pain, employed, time, days)
third.q.3 <- mutate(third.q.3, time.2 = round(time/days, 2)) # number of minutes per day spent between 50-74% of patient max
# Save to csv
tq <- third.q.3 %>%
    rename("minutes per week" = time, "days of recording" = days, "minutes per day" = time.2)
write.csv(tq, "../activity.plot/data/time.active.3q.csv", row.names = F)
# Ungroup
third.q.3 <- ungroup(third.q.3)

# Plot
# Univariate
qplot(x = pain, y = time.2, data = third.q.3, geom = "boxplot")
# Conditional (with employment status)
qplot(x = pain, y = time.2, data = third.q.3, facets = .~employed, colour = pain, geom = "boxplot")
# Wilcoxon rank sum test
with(third.q.3, wilcox.exact(time.2~pain, exact = T, paired = F))
# N, Median, IQR, range
foo <- third.q.3 %>%
    group_by(pain) %>%
    select(pain,time.2) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- third.q.3 %>%
    group_by(employed) %>%
    select(employed,time.2) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- third.q.3 %>%
    group_by(pain, employed) %>%
    select(pain, employed, time.2) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Second quartile (25-49%)
```{r second.q}
second.q <- group_by(data.10, id, date, pain, employed) # 25-49%
second.q.1 <- filter(second.q, pptmax.count<= 50 & pptmax.count>25)
second.q.2 <- ungroup(second.q.1)
second.q.2 <- group_by(second.q.2, id, pain, days, employed)
second.q.2 <- summarise(second.q.2, time = n())
second.q.2 <- group_by(second.q.2, id, pain, employed)
second.q.3 <- summarise(second.q.2, time = sum(time))
day <- tbl_df(data.frame(id.1 = as.numeric(as.character(fourth.q.3$id)), days = c(rep(7, 67), 5)))
id <- tbl_df(data.frame(id.1 = as.numeric(as.character(second.q.3$id))))
day.2 <- semi_join(day,id, by = "id.1")
second.q.3 <- bind_cols(second.q.3, day.2)
second.q.3 <- select(second.q.3, id, pain, employed, time, days)
second.q.3 <- mutate(second.q.3, time.2 = round(time/days, 2)) # number of minutes per day spent between 25-49% of patient max
# Save to csv
sq <- second.q.3 %>%
    rename("minutes per week" = time, "days of recording" = days, "minutes per day" = time.2)
write.csv(sq, "../activity.plot/data/time.active.2q.csv", row.names = F)
# Ungroup
second.q.3 <- ungroup(second.q.3)

# Plot
# Univariate
qplot(x = pain, y = time.2, data = second.q.3, geom = "boxplot")
# Conditional (with employment status)
qplot(x = pain, y = time.2, data = second.q.3, facets = .~employed, colour = pain, geom = "boxplot")
# Wilcoxon rank sum test
with(second.q.3, wilcox.exact(time.2~pain, exact = T, paired = F))
# N, Median, IQR, range
foo <- second.q.3 %>%
    group_by(pain) %>%
    select(pain,time.2) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- second.q.3 %>%
    group_by(employed) %>%
    select(employed,time.2) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- second.q.3 %>%
    group_by(pain, employed) %>%
    select(pain, employed, time.2) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### First quartile (0-24%)
```{r first.q}
first.q <- group_by(data.10, id, date, pain, employed) # 0-24%
first.q.1 <- filter(first.q, pptmax.count<= 25 & pptmax.count>0)
first.q.2 <- ungroup(first.q.1)
first.q.2 <- group_by(first.q.2, id, pain, days, employed)
first.q.2 <- summarise(first.q.2, time = n())
first.q.2 <- group_by(first.q.2, id, pain, employed)
first.q.3 <- summarise(first.q.2, time = sum(time))
day <- tbl_df(data.frame(id.1 = as.numeric(as.character(fourth.q.3$id)), days = c(rep(7, 67), 5)))
id <- tbl_df(data.frame(id.1 = as.numeric(as.character(first.q.3$id))))
day.2 <- semi_join(day,id, by = "id.1")
first.q.3 <- bind_cols(first.q.3, day.2)
first.q.3 <- select(first.q.3, id, pain, employed, time, days)
first.q.3 <- mutate(first.q.3, time.2 = round(time/days, 2)) # number of minutes per day spent between 0-24% of patient max
# Save to csv
fstq <- first.q.3 %>%
    rename("minutes per week" = time, "days of recording" = days, "minutes per day" = time.2)
write.csv(fstq, "../activity.plot/data/time.active.1q.csv", row.names = F)
# Ungroup
first.q.3 <- ungroup(first.q.3)

# Plot
# Univariate
qplot(x = pain, y = time.2, data = first.q.3, geom = "boxplot")
# Conditional (with employment status)
qplot(x = pain, y = time.2, data = first.q.3, facets = .~employed, colour = pain, geom = "boxplot")
# Wilcoxon rank sum test
with(first.q.3, wilcox.exact(time.2~pain, exact = T, paired = F))
# N, Median, IQR, range
foo <- first.q.3 %>%
    group_by(pain) %>%
    select(pain,time.2) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- first.q.3 %>%
    group_by(employed) %>%
    select(employed,time.2) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- first.q.3 %>%
    group_by(pain, employed) %>%
    select(pain, employed, time.2) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)
```

#### Time spent at 0 count
```{r zero}
zero.q <- group_by(data.10, id, date, pain, employed) #  = 0
zero.q.1 <- filter(zero.q, pptmax.count == 0)
zero.q.2 <- ungroup(zero.q.1)
zero.q.2 <- group_by(zero.q.2, id, pain, days, employed)
zero.q.2 <- summarise(zero.q.2, time = n())
zero.q.2 <- group_by(zero.q.2, id, pain, employed)
zero.q.3 <- summarise(zero.q.2, time = sum(time))
day <- tbl_df(data.frame(id.1 = as.numeric(as.character(fourth.q.3$id)), days = c(rep(7, 67), 5)))
id <- tbl_df(data.frame(id.1 = as.numeric(as.character(zero.q.3$id))))
day.2 <- semi_join(day,id, by = "id.1")
zero.q.3 <- bind_cols(zero.q.3, day.2)
zero.q.3 <- select(zero.q.3, id, pain, employed, time, days)
zero.q.3 <- mutate(zero.q.3, time.2 = round(time/days, 2)) # number of minutes per day spent at 0% of patient max
# Save to csv
zq <- zero.q.3 %>%
    rename("minutes per week" = time, "days of recording" = days, "minutes per day" = time.2)
write.csv(zq, "../activity.plot/data/time.active.zero.csv", row.names = F)
# Ungroup
zero.q.3 <- ungroup(zero.q.3)

# Plot
# Univariate
qplot(x = pain, y = time.2, data = zero.q.3, geom = "boxplot")
# Conditional (with employment status)
qplot(x = pain, y = time.2, data = zero.q.3, facets = .~employed, colour = pain, geom = "boxplot")
# Wilcoxon rank sum test
with(zero.q.3, wilcox.exact(time.2~pain, exact = T, paired = F))
# N, Median, IQR, range
foo <- zero.q.3 %>%
    group_by(pain) %>%
    select(pain,time.2) %>%
    rename(Group = pain) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- zero.q.3 %>%
    group_by(employed) %>%
    select(employed,time.2) %>%
    rename(Group = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)

foo <- zero.q.3 %>%
    group_by(pain, employed) %>%
    select(pain, employed, time.2) %>%
    rename(Group.Pain = pain, Group.Employed = employed) %>%
    summarise(N = n(), 
              Median = median(time.2), 
              IQR = IQR(time.2), 
              Min = min(time.2), 
              Max = max(time.2))
kable(foo, format = "markdown", align = "l", digits = 0)
```

## Session information
```{r SessionInfo}
sessionInfo()
```