---
title: 'Supplementary comparison of activity measures'
author: 'Peter Kamerman'
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
    md_document:
        variant: markdown_github
---

## Load required packages and set chunk options
```{r LoadPackages, include = FALSE}
# Load packages
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(coin)
library(ggplot2)
library(scales)
library(grid)
library(cowplot)

# Load palette
cb8.categorical <- c("#0072B2", "#D55E00", "#009E73", "#F0E442", "#56B4E9", "#E69F00", "#CC79A7", "#999999")


# knitr chunk options
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               fig.path = './figures/',
               fig.width = 11.7,
               fig.height = 8.2,
               dev = c('png', 'pdf'),
               tidy = FALSE, 
               tidy.opts = list(width.cutoff = 65))
```

## Process data
### Import
```{r DataImport, echo = TRUE}
data <- read_csv('./data/activity.bpi.eq5d.csv')
```

### Quick look
```{r QuickLook}
# Top-and-tail
head(data)
tail(data)
glimpse(data)
```

### Clean
```{r Clean, echo = TRUE}
data.1 <- data %>%
    mutate(ID = factor(ID),
           Pain = factor(Pain),
           eq5d_mobility = factor(eq5d_mobility,
                                  levels = c('I have no problems in walking about', 'I have some problems in walking about'),
                                  labels = c('I have no problems in walking about', 'I have some problems in walking about'),
                                  ordered = TRUE))
```

## Analyses
### Correlation between BPI interference with walking and time active
**(median time active daily over the past week)**
```{r BPIInterference.v.ActivityDuration}
# Extract required data
data.cor.duration <- data.1 %>%
    select(Pain, bpi_walk_interference, Median_activity_duration)

# Scatterplot 
scatter.plot.duration <- ggplot(data.cor.duration,
                 aes(x = bpi_walk_interference,
                     y = Median_activity_duration,
                     fill = Pain,
                     colour = Pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    labs(x = '\nMedian pain interference with walking (11-point NRS)',
         y = 'Median time active daily (minutes)\n',
         title = 'Pain interference vs time active\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.ticks = element_blank(), 
          panel.grid.major = element_line(colour = 'grey80'),
          panel.background = element_blank(),
          legend.position = 'none')
scatter.plot.duration

# Perform Spearman's correlation
result.cor.duration <- spearman_test(bpi_walk_interference ~ Median_activity_duration,
                            data = data.cor.duration)
result.cor.duration
```

### Correlation between BPI interference with walking and intensity of activity
**(median daily intensity of activity over the past week)**
```{r BPIInterference.v.ActivityIntensity}
# Extract required data
data.cor.intensity <- data.1 %>%
    select(Pain, bpi_walk_interference, Median_activity)

# Scatterplot 
scatter.plot.intensity <- ggplot(data.cor.intensity,
                 aes(x = bpi_walk_interference,
                     y = Median_activity,
                     fill = Pain,
                     colour = Pain)) +
    geom_jitter(size = 4, shape = 21, alpha = 0.5) +
    labs(x = '\nMedian pain interference with walking (11-point NRS)',
         y = 'Median daily intensity of activity (counts)\n',
         title = 'Pain interference vs intensity of activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.ticks = element_blank(), 
          panel.grid.major = element_line(colour = 'grey80'),
          panel.background = element_blank(),
          legend.position = 'none')
scatter.plot.intensity

# Perform Spearman's correlation
result.cor.intensity <- spearman_test(bpi_walk_interference ~ Median_activity,
                            data = data.cor.intensity)
result.cor.intensity
```

### EQ5D mobility rating vs time active
**(median time active daily over the past week)**
```{r EQ5DMobility.v.ActivityDuration}
# Extract required data
data.duration <- data.1 %>%
    select(Pain, eq5d_mobility, Median_activity_duration)

# Scatterplot 
duration.plot <- ggplot(data.duration,
                 aes(x = eq5d_mobility,
                     y = Median_activity_duration,
                     fill = eq5d_mobility,
                     colour = eq5d_mobility)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    labs(y = 'Median time active daily (minutes)\n',
         title = 'EQ5D mobility rating vs duration of activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    scale_x_discrete(breaks = c('I have no problems in walking about', 
                              'I have some problems in walking about'),
                   labels = c('I have no problems\nin walking about', 
                              'I have some problems\nin walking about')) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.title.x = element_blank(),
          axis.ticks = element_blank(), 
          panel.grid.major = element_line(colour = 'grey80'),
          panel.background = element_blank(),
          legend.position = 'none')
duration.plot

# Wilcoxon rank sum test
result.duration <- wilcox_test(Median_activity_duration ~ eq5d_mobility, 
                               data = data.duration)
result.duration
```

### EQ5D mobility rating vs intensity of activity
**(median daily intensity of activity over the past week)**
```{r EQ5DMobility.v.ActivityIntensity}
# Extract required data
data.intensity <- data.1 %>%
    select(Pain, eq5d_mobility, Median_activity)

# Scatterplot 
intensity.plot <- ggplot(data.intensity,
                 aes(x = eq5d_mobility,
                     y = Median_activity,
                     fill = eq5d_mobility,
                     colour = eq5d_mobility)) +
    geom_boxplot(fill = 'white', colour = 'white') +
    geom_boxplot(size = 1, alpha = 0.5) +
    geom_jitter() +
    labs(y = 'Median daily intensity of activity (counts)\n',
         title = 'EQ5D mobility rating vs intensity of activity\n') +
    scale_fill_manual(values = cb8.categorical) +
    scale_colour_manual(values = cb8.categorical) +
    scale_x_discrete(breaks = c('I have no problems in walking about', 
                              'I have some problems in walking about'),
                   labels = c('I have no problems\nin walking about', 
                              'I have some problems\nin walking about')) +
    theme_cowplot() +
    theme(plot.title = element_text(face = 'bold', 
                                    size = 18, 
                                    colour = 'black'),
          axis.text = element_text(size = 18, 
                                   colour = 'black'),
          axis.title = element_text(size = 18, 
                                    colour = 'black'),
          axis.title.x = element_blank(),
          axis.ticks = element_blank(), 
          panel.grid.major = element_line(colour = 'grey80'),
          panel.background = element_blank(),
          legend.position = 'none')
intensity.plot

# Wilcoxon rank sum test
result.intensity <- wilcox_test(Median_activity ~ eq5d_mobility, 
                               data = data.intensity)
result.intensity
```

## Session information
```{r SessionInfo}
sessionInfo()
```
